---
title       : Practical learning machine
subtitle    : Assignment Writeup - Prediction of how was Exercise Performed -
author      : Carlos Pinilla
job         : 
logo        : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow   # 
url:
  lib: ../../librariesNew
  assets: ../../assets
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---


```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F}
library(knitr); library(caret)
# make this an external chunk that can be included in any file
options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'fig/')

options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})
knit_hooks$set(plot = knitr:::hook_plot_html)
```


## Introduction

As nowdays we count with a large quantity of wereables, it is easy to collect a big amount of data relative to Human Activity Recognition. We count with a dataset with information about accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways and the goal is to use data to predict the exercises did by the volunters based in the data registered.

## Data

The training data for this project was taken from: 

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

The test data was taken from: 

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

---

## Getting the data

```{r loadData,cache=TRUE}
library(ggplot2); library(caret);
# Download data.
dir_data_training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
file_data_training <- "pml-training.csv"
download.file(url=dir_data_training, destfile=file_data_training)
dir_data_testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
file_data_testing <- "pml-testing.csv"
download.file(url=dir_data_testing, destfile=file_data_testing)

# Import the data treating empty values as NA.
training <- read.csv(file_data_training, na.strings=c("NA",""), header=TRUE)
colnames_train <- colnames(training)
testing <- read.csv(file_data_testing, na.strings=c("NA",""), header=TRUE)
colnames_test <- colnames(testing)
summary(training)
summary(testing)
```



---

## Get training/test sets

```{r trainingTest,dependson="loadData",cache=TRUE}
inTrain <- createDataPartition(training$classe, p=0.8, list=FALSE)
training_0 <- training[inTrain, ]
testing_0 <- training[-inTrain, ]
dim(Training_0); dim(Testing_0)
```



---

---

## Plot age versus wage


```{r ,dependson="trainingTest",fig.height=4,fig.width=6}
set.seed(408)
control<-trainControl(method="cv", number=5, allowParallel=T, verbose=T)
modelrf<-train(classe~.,data=training_0, method="rf", trControl=control, verbose=F)
modelrf$finalModel

predict_rf<-predict(modelrf, newdata=testing_0)
#we can see some statistics of the fit here:
confusionMatrix(predict_rf, testing_1$classe)

answers <- predict(rffit,testing);
answers
#testing$predRight <- pred==Subtesting$classe
#table(pred,testing$classe)
```


